---
title: "week-03-notes"
author: "Andrew Nalundasan"
date: "7/3/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(vtable)
library(jtools)
library(corrplot)

# load data and initial non-graphical exploration
data(mtcars)    # load data
head(mtcars)    # view top rows of dataset
vtable(mtcars)  # see mtcars in vtable
summary(mtcars) # summary statistics of mtcars
```
# Export Summs
```{r}
reg <- lm(mpg ~ cyl, mtcars)    # regress mpg on cyl
export_summs(reg)    # export summs table of reg
```
# R Graphics Cookbook: Correlation Matrix 

# Generate numerical correlation matrix

```{r}
mcor <- cor(mtcars, use="complete.obs")    # use="complete.obs" will get rid of NA's from dataset and remove NA's from matrix
# Print mcor and round to 2 digits
round(mcor, digits=2)
```

# plot the correlation matrix
```{r}
corrplot(mcor)
```
Figure: A correlation matrix

# Cosmetics to correlation matrix

```{r}
# matrix with colored squares and black labels, rotated 45 degrees along the top

corrplot(mcor, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45)
```
Figure: Correlation matrix with colored squares and black, rotated labels

# Fanciest correlation matrix

```{r}
# generate a lighter palette
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(mcor, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45,
         col = col(200), addCoef.col = "black", cl.pos = "n", order = "AOE")

# AOE - angular order of eigenvectors
```
Figure: Correlation matrix with correlation coefficients and no legend

