---
title: "week_03_notes"
author: "Andrew Nalundasan"
date: "7/3/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(vtable)
library(grid)
library(gridExtra)
library(lattice)
library(janitor)
library(GGally)
```

# Tidy EDA, Steps 1 and 2
- Analysis always begins with EDA - Exploratory Data Analysis
- Develop understanding of the data; look for mistakes in the data
- after a good EDA, you should have a good idea of what should be coming out of the analysis

**What am I looking for in EDA?**

1. things that are different that should not be
2. things that are similar that should not be
3. things that are relationships - variables that change in same or opposite directions (correlations)

**Base EDA is a 4 step process**

1. Uni-variate non-graphical

    - categorical (i.e. **factor**): count, proportion, %
    - quantitative: central tendancy, dispersion, skewness

2. Uni-variate graphical

    - categorical: barplot of factors
    - quantitative: histogram (try different bin), density, boxplot, quantile-normal plot

3. Multi-variate non-graphical

    - categorical: cross-tabulation if multiple categorical variables
    - quantitative: correlation (and covariance)

4. Multi-variate graphical
    
    - categorical: tile, mosaic graph, bar graph
    - quantitative: scatter plot, correlation plot
    - categorical and quantitative: distribution/summary by level of categorical variable, bargraphs, multiple boxplots, scatter plot with color/size/shape by factor or continuous variable

# Step 1 of EDA: Uni-variate non-graphical
```{r}
# load data
dis <- read_csv("Disability.csv")

# Recode date to age range, this is to fix an Excel formatting problem
dis <- dis %>% 
  mutate(age_group = recode(age_group, '12-Jun' = "6-12"))
```

## Non-graphical uni-variate analysis
Categorical and quantitative
```{r}
head(dis)
str(dis)
summary(dis)
vtable(dis)
```

+ Findings from Step 1

    - 1000 observations of 5 variables
    - 3 factor variables, 2 integer variables
    - need to change "12-Jun" to "6-12"

**Comments**

    - Age and expenditures likely right skewed so median better than mean 
    - Mostly White and Hispanic ethnicity
    - sample size of some ethnicity too small for statistically valid results
    - Gender is evenly mixed

**Question** 

    - Do expenditures vary substantially by age_group, ethnicity, or gender?

# Step 2.1 of EDA: Uni-variate graphical - categorical
- look at bar graphs of categorical variables to confirm non-graphical assumptions

**Comments**

    - many more obs age 21 and under
    - previous observations confirmed

**Question**

    - is ethnicity distribution representative of the population?

```{r}
# code univariate bar graphs using grid.arrange so can see all quant variables
grid.arrange(
  dis %>% 
    ggplot(mapping = aes(x = age_group)) + 
    geom_bar(), 
  dis %>% 
    ggplot(mapping = aes(x = ethnicity)) + 
    geom_bar(),
  dis %>% 
    ggplot(mapping = aes(x = gender)) + 
    geom_bar(),
  ncol = 2
)
```

# Step 2.2 of EDA: Uni-variate graphical - quantitative
- look at histograms and boxplots of quantitative variables to confirm non-graphical

**Comments**

    - It is unusual that age and expenditures have similar distributions
    - recipients younger than 25 and many expenditures below $18,000
    - don't throw outliers out just to save summary statistics
    - you must deal with outliers based on what you're trying to model
    - if focus is on demographic age below 50, then throw out outliers
    - if focus is on all age demographics, must keep outliers in

**Questions**

    - Does the skewed, lump pattern hold across ethnicity, age_group, and gender?
    - what should be done about outliers?

```{r}
# code histograms using grid.arrange so can see all quant variables together
grid.arrange(
  dis %>% 
    ggplot(mapping = aes(x = age)) + 
    geom_histogram(),
  dis %>% 
    ggplot(mapping = aes(x = expenditures)) + 
    geom_histogram(),
  # Add boxplots
  dis %>% 
    ggplot(mapping = aes(x = 1)) +
    geom_boxplot(mapping = aes(y = age)) + 
    coord_flip(),    # use to have same x-axis on both graphs
  dis %>% 
    ggplot(mapping = aes(x = 1)) + 
    geom_boxplot(mapping = aes(y = expenditures)) + 
    coord_flip(), 
  ncol = 2
)
```

# Step 3 of EDA: Multi-variate Non-Graphical Analysis

    - Non-graphical: *Essentially looking at tables*
    - with more experience, develop my own checklist for performing EDA depending on data I'm looking at
        - Better to make sure to get a full understanding of the dataset prior to diving into the particulars
    - **Don't get hung up on one question!**
    - multi-variate: relationships between variables
    - margins: provide totals of rows and total of columns
    - looking at tables can be limited, hence moving to visuals in Step 4

## Step 3.1 Multi-variate non-graphical - categorical

```{r}
# contingency/cross tabulation
dis %>% 
  tabyl(ethnicity, gender) %>%     # creates table of counts
  adorn_totals(where = c("row", "col")) %>%     # total margins
  kable()    # gives light formatting

```

    - Table of COUNTS with margins
        - total row and column are margins
        - bottom right cell shows total obs

```{r}
# proportion contingency/cross tabulation
dis %>% 
  tabyl(ethnicity, gender) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all") %>% 
  adorn_rounding(2) %>%     # round decimals
  kable()    # gives light formatting (from knitr)

```

    - Table of PROPORTIONS with margins
        - bottom right cell totals to 1
  
**Comments** 

    - confirms gender split is consistent across ethnicities
        - indicates we do not need to base our modeling on gender count
        - still need to look at age_group and expenditure by gender
    - low count ethnicities are outliers. Remove from dataset

**Questions**

    - need to check ethnicity/gender/expenditure balance
    - Can we get a larger sample?

## 3.1 - exploring the other variables

+ we want to look at **ALL** of these tables for all of our cross tabs and confirm **BALANCE**

```{r}
# relationship between age_group and gender
dis %>% 
  tabyl(age_group, gender) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  kable()
```

```{r}
# relationship between ethnicity and age_group
dis %>% 
  tabyl(ethnicity, age_group) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  kable()
```

**Comments**
    
    - age_group count is also split evenly by gender so don't need to model
    - ethnicities show similar pattern of increase **except Hispanic**

**Questions**

    - still need to check age_group/gender/expenditure **BALANCE**

## 3.1 - three way contingency table
```{r}
# relationship between ethnicity, age_group, gender
dis %>% 
  tabyl(ethnicity, age_group, gender) %>%    # Tables are categorized by LAST variable (gender)
  adorn_totals(where = c("row", "col"))
```

# 3.2 of EDA: Multi-variate non-graphical - Quantitative
## Correlations
```{r}
# Correlation table
dis %>% 
  select_if(is.numeric) %>%     # only selecting variables that are NUMERIC
  cor() %>% 
  round(2) %>% 
  kable()
```

**Comments**

    - large positive correlation between expenditure and age

**Questions**

    - Does positive correlation imply causation: individual expenditures increase with age?
    - never assume that high correlation means high causation

# Step 4.1 of EDA: Multi-variate - graphical - categorical

+ looking for relationships between variables

## 4.1 Multi-variate graphical - categorical

```{r}
# code univariate bar graphs using grid.arrange so can see graphs together
grid.arrange(
  
  dis %>% 
    ggplot(mapping = aes(x = ethnicity, fill = age_group)) + 
    geom_bar(position = "dodge"), 
  
  dis %>% 
    ggplot(mapping = aes(x = age_group, fill = gender)) + 
    geom_bar(position = "dodge"),
  
  dis %>% 
    ggplot(mapping = aes(x = ethnicity, fill = gender)) +
    geom_bar(position = "dodge"),
  
  ncol = 2
)
```

**Comments**

    - there appears to be a difference in age_group across ethnicities
    - need to take a closer look
        - gender is closely balanced by ethnicity and age_group
  
**Questions**

    - What does it mean if there are differences by age_group across ethnicities?

## 4.1 - alternative view of categorical variables
- 2 ways to view the same graph
- flip the axis to be easier to read
- interesting to see it because it's based off **PROPORTIONS**

```{r}
# more detail on ethnicity and age
grid.arrange(
  # Cluster of counts
  dis %>% 
    ggplot(mapping = aes(x = ethnicity, fill = age_group)) + 
    geom_bar(position = "dodge") + 
    coord_flip(),
  
  dis %>% 
    ggplot(mapping = aes(x = ethnicity, fill = age_group)) +
    geom_bar(position = "fill") + 
    coord_flip(),
  
  ncol = 1
  
)
```

**Comments**

+ Asian, Black, and White have similar age distributions
    
    - Hispanic and Multi-Race have low count for older than 22
    - recall, cannot make valid statements about the other 3 ethnicities

**Question**

    - How do median expenditures vary for the ethnicity and age_groups?
    - Need larger sample to determine if low count for older than 22 in Hispanic and Multi Race

# 4.1 - alternative alternative way to view categorical variables
- tile or mosaic plots show multiple categorical variables 

```{r}
# Note - have to calculate count and provide a variable for filling the tiles
dis %>% 
  group_by(age_group, ethnicity) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(age_group, ethnicity)) +
  geom_tile(aes(fill = count))    # default shows darker being lower count
        # use fill = -count to flip the color scheme
```

**Comments**

    - Easy to see difference between Hispanic and White
        - Asian, Black, and Multi-Race harder to see difference due to lower count
  
# 4.2 Multi-variate graphical - quantitative
- most common is to use scatter plot
- we only have 2 quantitative variables, so can compare those in 1 visual

```{r}
# flipped tile scales
dis %>% 
  group_by(age_group, ethnicity) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(age_group, ethnicity)) +
  geom_tile(aes(fill = -count))    # default shows darker being lower count
```

- Problem with this is that the scale is different from the original. But lighter is less and darker is more

```{r}
# code scatterplots using grid.arrange so can see all quant variables of interest
dis %>% 
  ggplot(mapping = aes(x = age, y = expenditures)) + 
  geom_point()
```

**Comments**

+ Interesting pattern
  
    - need to capture distinct groups
    - 17 under, 18-21, and 22 up are 3 different groups

**Questions**

    - Is the 3 group age/expenditures pattern the same across ethnicities?

# Final Step of EDA: Multi-variate graphical - categorical and quantitative
- Interplay between Categorical and Quantitative Variables
- summary statistics of quantitative variables by the factor variables
1. Scatter plot: with or without "trend" line

    - quantitative variable on each axis
    - factor variable levels differentiated by color, size, shape, etc
2. Box plot
  
    - categorical/factor variable levels on one axis
    - box plot of quantitative variable on other axis
3. Bar graphs: clustered, stacked, proportion
  
    - categorical/factor variable levels on one axis
    - summary of quantitative variable on other axis
        - sum, mean, median, count, etc
    - additional factor variable levels with fill()
4. Heat map: aka mosaic, tile, raster
  
    - factor variable on each axis
    - summary of quantitative variable by shade/color in legend
        - sum, mean, median, count, etc
    
```{r}
# high level overview of the dataset. 
# this is like graphical summary statistics
dis %>% 
  ggpairs()
```

**Comments**

    - gender is balanced looking both directins on correlation graph

**Questions**

    - Look more closely at: ethnicity vs. age, ethnicity vs. expenditures

# Step 4.3 - multi-variate graphical - detailed scatter plot

```{r}
# show code
grid.arrange(
  # age, expenditures, and ethnicity
  dis %>% 
    ggplot(mapping = aes(x = age, y = expenditures, color = ethnicity)) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE), 
  
  # age, expenditures, and gender
  dis %>% 
    ggplot(mapping = aes(x = age, y = expenditures, color = gender)) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE), 
  
  ncol = 1
)
```

**Comments**

    - Similar relationship between age and expenditures across ethnicities and gender
        - similar slope lines indicate this
        - again see no differences by gender
  
**Questions** 
    
    - Look at the age/expenditure/ethnicity with just large sample ethnicities
        - try faceted scatter plot

## Same data from scatter plots as boxplots

```{r}
# show code
grid.arrange(
  # age, expenditures, and ethnicity
  dis %>% 
    ggplot(mapping = aes(x = ethnicity, y = age)) + 
    geom_boxplot(), 
  # age, expenditures, and gender
  dis %>% 
    ggplot(mapping = aes(x = ethnicity, y = expenditures)) + 
    geom_boxplot(), 
  
  ncol = 1
)
```

**Comments**

+ Ethnicity vs age
    
    - multi-race has lowest age distribution, small sample n = 26
    - hispanic has slightly lower age distribution than other large groups

+ Ethnicity vs Expenditures
    
    - Hispanic and Multi-Race have lowest dstribution of expenditures
    - could be driven (caused) by age distribution since younger receive less expenditures
    - there are so many younger Hispanic and small expenditures that older and larger expenditures are relative outliers
  
**Questions**

    - Why is the distribution of Hispanic and Multi-Race younger than others?

## Same data from scatter plots and boxplots as bar graphs

```{r}
# group the median expenditure and age in same graph
grid.arrange(
  # Median expenditures by ethnicity and gender
  dis %>%
    group_by(ethnicity, gender) %>% 
    summarise(med_exp_000 = median(expenditures) / 1000) %>% 
    ggplot(mapping = aes(x = ethnicity, y = med_exp_000, fill = gender)) + 
    geom_bar(stat = "identity", position = "dodge") +
    coord_flip(),
  
  # median expenditures by ethnicity and gender
  dis %>% 
    group_by(ethnicity, gender) %>% 
    summarise(med_age = median(age)) %>% 
    ggplot(mapping = aes(x = ethnicity, y = med_age, fill = gender)) + 
    geom_bar(stat = "identity", position = "dodge") + 
    coord_flip(), 
  
  ncol = 1
)
```

**Comments**

    - Median expenditures substantially higher for White Male
        - Black Male higher than Black Female

**Questions**

    - Need to examine why median expenditures so much more for White Male thatn White Female


## Same data from scatter plots and boxplots and bar graphs as mosaic

```{r}
#Note have to calculate and provide a variable for filling the graph
dis %>% 
  group_by(age_group, ethnicity) %>% 
  summarise(med_exp_000 = median(expenditures) / 1000) %>% 
  ggplot(aes(age_group, ethnicity)) + 
  geom_tile(aes(fill = med_exp_000))
```

**Comments**

    - Factor variable on each axis
    - expenditures get higher with age

- Base EDA is done to setup the analysis. Foundation required for success of the model and project!

# Detailed EDA
- Answer questions raised in Base EDA

3 types of EDA's: 
1. Base EDA - raise questions and learn about data
2. Detailed EDA - answer and address these questions
3. Statistical EDA - learn how valid the findings are that we found

    - modeling and other things to better understand what's going on

2 Groups of questions: 
1. Questions to be answered outside the data

    - don't let the data limit your thought process
    - can't move outliers that just improves model. must be intentional with these decisions

2. Questions to be answered by further analysis of the data

# Facet Graph
- create individual graph per factor level

```{r}
# create vector of ethnicities of interest (remove small sample ethnicities)
eth_ana <- c("Asian", "Black", "Hispanic", "Multi-Race", "White")

# Faceted histogram without small sample ethnicities
# facet show individual graph by factor level
dis %>% 
  filter(ethnicity %in% eth_ana) %>% 
  ggplot(mapping = aes(x = expenditures)) + 
  geom_histogram() + 
  facet_wrap( ~ ethnicity)

# faceted histogram without small sample age_group
dis %>% 
  filter(ethnicity %in% eth_ana) %>% 
  ggplot(mapping = aes(x = expenditures)) +
  geom_histogram() + 
  facet_wrap( ~ age_group)

```

## Are expenditures similar by age_group, ethnicity, or gender?

```{r}
# faceted scatter plot by ethnicity
dis %>% 
  filter(ethnicity %in% eth_ana) %>%   # "%in%" choose whatever is in 'ethnicity' that is also in 'eth_ana'
  ggplot(mapping = aes(x = age, y = expenditures)) + 
  geom_point() + 
  facet_wrap( ~ ethnicity)    # "~" spreads out the specified variable
```

**Comment**

    - yes, the pattern is consistent across ethnicities

## Do expenditures differ substantially by age_group, ethnicity, or gender?

```{r}
# age/expenditures relationship differ by age_group
dis %>% 
  filter(ethnicity %in% eth_ana) %>% 
  ggplot(mapping = aes(x = age, y = expenditures)) + 
  geom_point() + geom_smooth(mapping = aes(color = age_group), method = "lm", se = FALSE)
```

**Comment**

    - Under 18 age group all act similar

## Why are median expenditures more for White Male than White Female?
```{r}
# different views of White ethnicity by gender
grid.arrange(
  # Boxplots of expenditures
  dis %>% 
    filter(ethnicity == "White") %>% 
    ggplot(mapping = aes(x = gender, y = expenditures)) + 
    geom_boxplot(),
  
  # Density of expenditures
  dis %>% 
    filter(ethnicity == "White") %>% 
    ggplot(mapping = aes(x = expenditures, fill = gender)) + 
    geom_density(alpha = 0.5),
  
  # Expenditures count
  dis %>% 
    filter(ethnicity == "White") %>% 
    group_by(age_group, gender) %>% 
    summarize(n = n()) %>% 
    ggplot(mapping = aes(x = age_group, y = n, fill = gender)) + 
    geom_bar(stat = "identity", position = "dodge"),
  
  # Median Expenditure
  dis %>% 
    filter(ethnicity == "White") %>%
    group_by(age_group, gender) %>% 
    summarize(med_exp_000 = median(expenditures) / 1000) %>% 
    ggplot(mapping = aes(x = age_group, y = med_exp_000, fill = gender)) + 
    geom_bar(stat = "identity", position = "dodge"),
  
  ncol = 2
    
)
```

**Comments**
    
    - no bias found for White Males. Female expenditures are higher in age_groups except for 22-50. 
    - Detailed EDA will lead to modeling


# Facets
- separate graphs for each factor level
- not a geom()
- different from grid.arrange()
    - used when trying to show different graphs in the same plot (grouping them together)
  
- facets are different because they're the same graphs but for different factor levels
    - same axes for all graphs
    - will provide as many graphs as there are factor levels
    - great for making comparisons
    
## Facet wraps

```{r}
# create facet graph
dis %>% 
  group_by(ethnicity) %>% 
  ggplot(mapping = aes(x = age, y = expenditures)) + 
  geom_point() + 
  facet_wrap(~ ethnicity)
```
    
```{r}
# control columns of facet graph
dis %>% 
  group_by(ethnicity) %>% 
  ggplot(mapping = aes(x = age, y = expenditures)) + 
  geom_point() + 
  facet_wrap(~ ethnicity, ncol = 2)
```

## Facet grids
- look at facets divided by ethnicity and gender
- axes tick labels are only on left and bottom

```{r}
# Create facet grid to examine differencees by Region, Category, and Segment
dis %>% 
  group_by(ethnicity, gender) %>% 
  ggplot(mapping = aes(x = age, y = expenditures)) + 
  geom_point() + 
  facet_grid(ethnicity ~ gender)
```

**comments**

    - these don't differ much by gender
    
## Facet grid by different geoms
- bar and line graphs MUST have summary of data somewhere


```{r}
# create facet grid to examine differences by Region, Category, and Seg
dis %>% 
  group_by(age_group, ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = age_group, y = med_exp, fill = age_group)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  facet_grid(ethnicity ~ gender)
```

```{r}
# facet grid line graph
dis %>% 
  group_by(age_group, ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = age_group, y = med_exp, color = ethnicity)) + 
  geom_line(mapping = aes(group = ethnicity)) + 
  facet_grid(ethnicity ~ gender)
```

# Line Graphs

+ Line graphs imply motion

    - do not use to connect factor variables (gender, color...why connect these?)
    - possible to connect ordered factors
    
+ Most often used with time series

    - comparisons over time
    - summary over time
        - median expenditure per time perios
        
+ Key: Implied continuity along x-axis

    - some kind of motion
    - like a scatter plot, where causation related to time
    
+ Must have a time component (date variable)

```{r}
# Create random dates for making line graphs
dis$date <- sample(seq(as.Date("2015/01/01"), as.Date("2018/01/01"), by = "month"), 1000, replace = TRUE)

# View data with date
summary(dis)
```

```{r}
# Look at expenditures over time
ggplot(data = dis, mapping = aes(x = date, y = expenditures)) + 
  geom_line()
```

**Comments**

    - all expenditures are being measured on a specific month
    - we're getting obs that the line graph is trying to connect on the same date, hence VERTICLE Line
    - Instead could show summary statistic for each month
    
```{r}
# look at median expenditures over time
dis %>% 
  group_by(date) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = date, y = med_exp)) + 
  geom_line()
```

**Comments**

    - show quantitative varibales over time
    - could show the count of a factor over time
    - must make a quantitative variable related to the factor
    
## Median expenditures over time by age_group

```{r}
# Look at median expenditures over time
dis %>% 
  group_by(date, age_group) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = date, y = med_exp, linetype = age_group)) + 
  geom_line()
  
```

- Note: color = age_group differentiates factor levels

```{r}
# Look at median expenditures over time
dis %>% 
  group_by(date, age_group) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = date, y = med_exp, color = age_group)) + 
  geom_line()
```

## Look at all factor variables

```{r}
# use grid.arrange to put graphs together
grid.arrange(
  # graph mean annual profit over time by Region
  dis %>% 
    group_by(date, age_group) %>% 
    summarize(med_exp = median(expenditures)) %>% 
    ggplot(mapping = aes(x = date, y = med_exp, color = age_group)) + 
    geom_line(), 
  
  # graph mean annual profit over time by Category
  dis %>% 
    group_by(date, ethnicity) %>% 
    summarize(med_exp = median(expenditures)) %>% 
    ggplot(mapping = aes(x = date, y = med_exp, color = ethnicity)) + 
    geom_line(), 
  
  # graph mean annual profit over time by Segment
    dis %>% 
    group_by(date, gender) %>% 
    summarize(med_exp = median(expenditures)) %>% 
    ggplot(mapping = aes(x = date, y = med_exp, color = gender)) + 
    geom_line(), 
  
  ncol = 2
)
```

**Comments**

    - if using black and white printing, need to do "linestyle = x" rather than "color = x"
    - avoid spaghetti graph
    - avoid squishing lines
    

## with ordered factor variables

```{r}
# Look at median expenditures by age_group and ethnicity
dis %>% 
  group_by(age_group, ethnicity) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = age_group, y = med_exp, color = ethnicity)) + 
  geom_line(mapping = aes(group = ethnicity)) 
```

+ change over time where the quantity really means something
