---
title: 'Quick Project #1'
author: "Andrew Nalundasan"
date: "7/8/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r}
# Course: OMSBA 5210
# Purpose: Quick Project # 1
# Date: July 8, 2021
# Author: Andrew Nalundasan
# Pushing commits: 7/21/2021
```

# Load Libraries

```{r}
# The tidyverse package contains ggplot2, dplyr, and several other packages
library(tidyverse)

# The gridExtra package contains grid.arrange function used to combine plots in the same window
library(gridExtra)

# The janitor package contains tidyverse functions for cross-tables
library(janitor)

# The knitr package contains some table formatting functions
library(knitr)

# The GGally package contains a custom correlation plot
library(GGally)

# vtable for reppin NHK
library(vtable)

# correlation matrix
library(corrplot)

# to get $ and , on axis in graph
library(scales)

```

# Load data

```{r}
# Load and prepare data 
zillow <- read_csv("../03_raw_data/qp1_data.csv")

# look at the data
zillow
```

# Base EDA Step 1: Uni-variate non-graphical EDA

```{r}
# top of the dataset
head(zillow)
```

+ data appears to be tidy

    - data in each column is of the same variable type
    - no duplicate columns
    - each row is an observation for the sale of a specific home

```{r}
# check the structure of the data
vtable(zillow)
```

+ 3 continuous integer variables

    - price, sqft_living, sqft_lot
    
+ factor variables

    - waterfront: binary yes/no
    - view: 5 ordered levels
    - grade: 13 ordered levels 
    - condition: 5 ordered levels
    - bedrooms: outlier of 33?
    - bathrooms: outlier of 8?
    - floors

```{r}
# check descriptive statistics of the data
summary(zillow)
```

**Comments**

    - mean home price is 0.5+ million. sounds about right
        - appear to be right-skewed because mean >> median
            - median home price will be more representative than mean
        - 7.7 MM max is really high compared to mean of 0.5 MM. 
  
    - is max 33 bedrooms an outlier?
        - much higher than 3rd quartile of 4
        
    - is max 8 bathrooms an outlier?
        - much higher than 3rd quartile of 2.5
        
    - is max 13540 sqft_living an outlier?
        - appear to be right-skewed because mean >> median
            - median more representative than mean
        - much higher than 3rd quartile of 2550
        
    - is max 1651359 sqft_lot an outlier?
        - appear to be right-skewed because mean >> median
            - median more representative than mean
        - much higher than 3rd quartile of 10688
        
    - floors seem evenly distributed
    
    - waterfront is binary yes/no
    
    - seems like not many houses have a nice view
    
    - is max 13 grade an outlier?
        - much higher than 3rd quartile of 8
        
    - condition seems evenly distributed
    
    - is 1900 yr_built an outlier?
        - much older than 1st quartile of 1951 and mean of 1971
        
    - yr_renovated looks odd
    
    
**Questions**

+ What's the relationship between price and # bedrooms, # bathrooms, sqft_living/sqft_lot, # floors?

+ What's the relationship between price and grade or condition?

+ what's the relationship between price and zip_code?

+ MAIN QUESTION for me: What drives price?


# Base EDA Step 2: Uni-variate graphical EDA

- Examine each variable individually
- Examine how many observations in each variable

## Factor variables

- show all factor variables together

```{r}
# Only use bar graphs (like histogram) because nothing like boxplot for cateorical variables
grid.arrange(

# waterfront
ggplot(data = zillow, mapping = aes(x = waterfront)) +
  geom_bar(),

# view
ggplot(data = zillow, mapping = aes(x = view)) +
  geom_bar(), 

# grade
ggplot(data = zillow, mapping = aes(x = grade)) +
  geom_bar(),

# condition
ggplot(data = zillow, mapping = aes(x = condition)) + 
  geom_bar(),

ncol = 2 )
```

**Comments**

    - Waterfront property is rare
    - Homes with a view better than 0 are rare
    - Most homes are grade 7-8
    - Most homes have condition of 3
    - Most homes have 3 bedrooms
    - Most homes have 2.5 bathrooms or less
    
**Questions**

+ Is price distribution representative of these variables?

## Other variables

- show all other variables together

```{r}
# Only use bar graphs (like histogram) because nothing like boxplot for categorical variables
grid.arrange(

# bedrooms
zillow %>% 
  ggplot(mapping = aes(x = bedrooms)) + 
  geom_bar(),

# bathrooms
zillow %>% 
  ggplot(mapping = aes(x = bathrooms)) + 
  geom_bar(),

# sqft_living
zillow %>% 
  ggplot(mapping = aes(x = sqft_living)) +
  geom_bar(),

# floors
ggplot(data = zillow, mapping = aes(x = floors)) +
  geom_bar(),

# zip_code
ggplot(data = zillow, mapping = aes(x = zip_code)) +
  geom_bar(),

# yr_built
ggplot(data = zillow, mapping = aes(x = yr_built)) +
  geom_bar(),

# yr_renovated
zillow %>% 
  filter(yr_renovated > 1925) %>% 
  ggplot(mapping = aes(x = yr_renovated)) + 
  geom_bar(),

ncol = 2)
```

**Comments**

    - record with 33 bedrooms is throwing off my data
    - records with 4+ bathrooms throwing off my data
    - records with 5000+ sqft_living throwing off my data
    - most homes have either 1 or 2 floors
        - split levels are rare
        - 3 floors are rare
    - not sure what to do about zip_codes
    - big dip during great depression (30s) and WWII (40s) for homes built
        - boom in houses built in 50s post war, along with Baby Boomer generation
        - assumption that home building slows down during war times
        - Vietnam War in 60s
        - Cold War/Gulf War in 80s - 90s?
    - housing bubble burst in 00s
    - home renovations increase from 1980 onward
    
**Questions**

+ Is price distribution representative of these variables?

+ How does price fluctuate with yr_built?

```{r}
# Removing outliers
grid.arrange(

# bedrooms
  zillow %>% 
    filter(bedrooms < 5) %>% 
    ggplot(mapping = aes(x = bedrooms)) +
    geom_bar(),
  
  # bathrooms
  zillow %>% 
    filter(bathrooms < 4) %>% 
    ggplot(mapping = aes(x = bathrooms)) + 
    geom_bar(), 
  
  # sqft_living
  zillow %>% 
    filter(sqft_living < 4000) %>% 
    ggplot(mapping = aes(x = sqft_living)) +
    geom_bar(),
  
  # yr_renovated
  zillow %>% 
    filter(yr_renovated > 1975) %>% 
    ggplot(mapping = aes( x = yr_renovated)) + 
    geom_bar(),

  ncol = 1)

```

**Comments**

    - filtered bedrooms to show < 5
    - filtered bathrooms to show < 4
    - filtered sqft_living to show < 4000

## Quantitative variables

### Price

```{r}
# price
grid.arrange(
  # add histogram
  ggplot(data = zillow, mapping = aes(x = price)) +
    geom_histogram() + 
    scale_x_continuous(labels = dollar), 
  
  # add boxplot
  ggplot(data = zillow, mapping = aes(x = 1)) + 
    geom_boxplot(mapping = aes(y = price)) + 
    scale_y_continuous(labels = dollar) + 
    coord_flip()
)
```

**Comments**

    - Confirms there is a skewed distribution
        - Use median rather than mean
        - 2+MM are all outliers
    
**Questions**

+ Does price differ substantially by my factor variables?

+ should I remove 2+ MM from dataset?

```{r}
# price with outliers removed
grid.arrange(
  # add histogram
  zillow %>% 
    filter(price < 2000000) %>% 
    ggplot(mapping = aes(x = price)) +
    geom_histogram() + 
    scale_x_continuous(labels = dollar), 

  # add boxplot
  zillow %>%
    filter(price < 2000000) %>% 
    ggplot(mapping = aes(x = 1)) + 
    geom_boxplot(mapping = aes(y = price)) + 
    scale_y_continuous(labels = dollar) + 
    coord_flip()
)
```

**Comments**

    - Median price is 450k
    
**Questions**

+ When were the houses built where price <= 450k?

### Square Foot Living

```{r}
# sqft_living
grid.arrange(
  # add histogram
  zillow %>% 
    filter(sqft_living < 4000) %>% 
    ggplot(mapping = aes(x = sqft_living)) +
    geom_histogram(), 
  
  # add boxplot
  zillow %>% 
    filter(sqft_living < 4000) %>% 
    ggplot(mapping = aes(x = 1)) + 
    geom_boxplot(mapping = aes(y = sqft_living)) + 
    coord_flip()
)
```

**Comments**

    - Confirms there is a skewed distribution
    - majority of homes < 4000 sqft_living
        - kept filter consistent to show < 4000 sqft_living
    
**Questions**

+ Does sqft_living differ substantially by my factor variables?

+ I wonder how price is distributed according to sqft_living?

### Square Foot Lot

```{r}
# sqft_lot
grid.arrange(
  # add histogram
  zillow %>% 
  filter(sqft_lot < 10000) %>% 
  ggplot(mapping = aes(x = sqft_lot)) +
    geom_histogram(), 
  
  # add boxplot
  zillow %>% 
  filter(sqft_lot < 10000) %>% 
  ggplot(mapping = aes(x = 1)) + 
    geom_boxplot(mapping = aes(y = sqft_lot)) + 
    coord_flip()
)
```

**Comments**

    - Left-skewed distribution
    
**Questions**

+ How is price distributed according to sqft_lot?


# Base EDA Step 3: Multi-variate non-graphical

## Categorical 

```{r}
# Contingency/cross table

zillow %>% 
  tabyl(grade, waterfront) %>%     # creates table of counts
  adorn_totals(where = c("row", "col")) %>%    # Total margins
  kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped")

```


```{r}
# Proportion contingency/cross table
zillow %>% 
  tabyl(grade, waterfront) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all") %>% # creates proportions
  adorn_rounding(2) %>%  # round decimals
  kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped")
```


**Comments**

    - 99% of homes do NOT have waterfront view
    - Most number of waterfront homes are grade 8 (40 obs)
    - grade 7 and grade 8 non-waterfront are majority of homes
    - low number of homes grade 11+
    
    
**Questions**

+ how important is waterfront property?

    - Decision to remove waterfront from analysis

```{r}
zillow %>% 
  tabyl(grade, view) %>%     # creates table of counts
  adorn_totals(where = c("row", "col")) %>%    # Total margins
  kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped")
  
```

```{r}
# Proportion contingency/cross table
zillow %>% 
  tabyl(grade, view) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all") %>% # creates proportions
  adorn_rounding(2) %>%  # round decimals
  kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped")
```

**Comments**

    - 70% of homes rated grade 7 and grade 8
    - 90% of homes do not have a view (view = 0)
        - besides homes with no view, homes with view = 2 has the most (4%)
        
**Questions**

+ Why don't these homes have views?

+ When were the homes with view = 0 built?

```{r}
zillow %>% 
  tabyl(grade, condition) %>%     # creates table of counts
  adorn_totals(where = c("row", "col")) %>%    # Total margins
  kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped")
```


```{r}
# Proportion contingency/cross table
zillow %>% 
  tabyl(grade, condition) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all") %>% # creates proportions
  adorn_rounding(2) %>%  # round decimals
  kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped")
  
```

**Comments**

    - 65% of homes are condition = 3
        - condition = 4 is next highest at 26%

**Questions**

+ How does price differ between condition = 3 and condition = 4?

## Quantitative

```{r}
# Correlation table
mcor <- cor(zillow, use="complete.obs")    # use="complete.obs" will get rid of NA's from dataset and remove NA's from matrix

# Print mcor and round to 2 digits
round(mcor, digits=2)
```

**Comments**

    - so many variables to look at, need to see this in a different visual

```{r}
# correlation matrix
corrplot(mcor)
```

**Comments**

    - Colors are hard to match up. Still difficult to interpret

```{r}
# updates to correlation matrix
corrplot(mcor, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45)
```

**Comments**

    - sqft_lot not correlated with anything
    - price highly correlated with bathrooms, sqft_living, grade
    - sqft_living highly correlated with price, bathrooms, grade
    - zip_code has lower correlation on price than I assumed
        - mostly negative correlations
    - bathrooms have higher correlation with price than bedrooms
        
+ variables of interest

    - price on:
      - bathrooms: 0.53
      - sqft_living: 0.70
      - grade: 0.67

**Questions**

+ How does price distribution look between bathrooms, sqft_living, and grade?

```{r}
# generate a lighter palette
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(mcor, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45,
         col = col(200), addCoef.col = "black", cl.pos = "n", order = "AOE")
```

**Comments**

    - colors and numbers work, but too squished. 
    - will use other correlation matrix from above
    

# Base EDA Step 4: Multi-variate graphical

## Step 4.1: Categorical

- Bar graphs with multiple categorical variables
- Focus on grade and bathrooms as main categorical variables

```{r}
# Code univariate bar graphs using grid.arrange so can see graphs together

# Create vector of bathrooms want to analyze
bath <-  c("1", "1.5", "2", "2.5", "3", "3.5", "4")
bath <- as.factor(bath)

grid.arrange(
  
  zillow %>% 
    mutate(condition = as.factor(condition)) %>% 
    ggplot(mapping = aes(x = grade, fill = condition)) + 
    geom_bar(position = "dodge"),
  
  zillow %>% 
    mutate(view = as.factor(view)) %>%
    ggplot(mapping = aes(x = grade, fill = view)) + 
    geom_bar(position = "dodge"),
  
  zillow %>%
    filter(bathrooms %in% bath) %>% 
    mutate(bathrooms = as.factor(bathrooms)) %>% 
    ggplot(mapping = aes(x = grade, fill = bathrooms)) + 
    geom_bar(position = "dodge"), 
  
  ncol = 2
)


```

**Comments**

    - needed to mutate numerical variables as factors
    - confirms knowledge about condition, and view
        - homes in great condition (5) are low
    - grades 7 and 8 have most observations
        - most homes have either 1 or 2.5 bathrooms

**Questions**

+ are the condition = 5 homes brand new built homes or newly renovated homes?

## Categorical - condition/grade

```{r}
# More detail on condition and grade
grid.arrange(
  # Cluster of counts
  zillow %>% 
    mutate(condition = as.factor(condition)) %>% 
    ggplot(mapping = aes(x = grade, fill = condition)) + 
    geom_bar(position = "dodge") + 
    coord_flip(), 
  
  # proportion of counts
  zillow %>% 
    mutate(condition = as.factor(condition)) %>% 
    ggplot(mapping = aes(x = grade, fill = condition)) + 
    geom_bar(position = "fill") + 
    coord_flip()
)
```

**Comments**

    - condition = 4 is consistent across grades 5-7
    - condition = 3 is is majority of most grades
        - grade 3 condition has the most homes in top condition
    - all grade 0 homes have condition = 1
    
**Questions** 

+ how does price vary by grade and condition

```{r}
# Tile graph - need count
zillow %>% 
  group_by(condition, grade) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(condition, grade)) + 
  geom_tile(aes(fill = -count))
```

**Comments**

    - grades 7-8 mostly condition = 3

## Step 4.2: Quantitative

```{r}
# Code scatterplots using grid.arrange so can see all quant variables together
grid.arrange(
  # price on bathrooms
  zillow %>% 
    filter(bathrooms < 4) %>% 
    ggplot(mapping = aes(x = bathrooms, y = price)) + 
    geom_point(position = "jitter") + 
    scale_y_continuous(labels = dollar), 
  
  # price on bedrooms
  zillow %>% 
    filter(bedrooms < 5) %>% 
    ggplot(mapping = aes(x = bedrooms, y = price)) + 
    geom_point(position = "jitter") + 
    scale_y_continuous(labels = dollar),
  
  # price on sqft_living
  zillow %>% 
    filter(sqft_living < 4000) %>% 
    ggplot(mapping = aes(x = sqft_living, y = price)) + 
    geom_point(position = "jitter") +    # can't see points all on a single line. Jitter helps see multiple points on same point
    scale_y_continuous(labels = dollar),
  
  # price on grade
  zillow %>% 
    ggplot(mapping = aes(x = grade, y = price)) + 
    geom_point(position = "jitter") + 
    scale_y_continuous(labels = dollar), 
  
  # price on condition
  zillow %>% 
    ggplot(mapping = aes(x = condition, y = price)) + 
    geom_point(position = "jitter") + 
    scale_y_continuous(labels = dollar), 
  
  # price on yr_built
  zillow %>% 
    ggplot(mapping = aes(x = yr_built, y = price)) + 
    geom_point(position = "jitter") + 
    scale_y_continuous(labels = dollar), 
  
  # price on yr_renovated
  zillow %>% 
    filter(yr_renovated > 1990) %>%   # removed outliers
    ggplot(mapping = aes(x = yr_renovated, y = price)) + 
    geom_point(position = "jitter") + 
    scale_y_continuous(labels = dollar), 
  
  # price on zip_code
  zillow %>% 
    ggplot(mapping = aes(x = zip_code, y = price)) + 
    geom_point(position = "jitter") + 
    scale_y_continuous(labels = dollar), 
  
  ncol = 2
)


```

**Comments**

    - price positively correlated with most variables, this makes sense
        - more bathrooms, higher price
        - more bedrooms, higher price
        - more sqft_living, higher price
    - price higher at condition = 3 and condition = 4
        - condition = 5 does not guarantee high price
    - price does not increase with more recent renovations
        - filtered yr_renovated to focus on 1990 onward
    - price not effected much by yr_built
    
**Questions**

+ Why is price higher at condition = 3 and condition = 4 than condition = 5?

    - perhaps because count of condition = 5 is lower?
    
+ What is relationship between grade and yr_renovated?

    - are house flippers getting sloppy?
    
## Step 4.3: Categorical and quantitative

```{r}
grid.arrange(
  # condition, price, grade
  zillow %>% 
    mutate(condition = as.factor(condition)) %>% 
    ggplot(mapping = aes(x = grade, y = price, color = condition)) + 
    geom_point(position = "jitter") + 
    scale_y_continuous(labels = dollar) + 
    geom_smooth(method = "lm", se = FALSE), 
  
  # price, bedrooms, bathrooms
  zillow %>% 
    filter(bedrooms < 5) %>% 
    filter(bathrooms %in% bath) %>% 
    mutate(bathrooms = as.factor(bathrooms)) %>% 
    ggplot(mapping = aes(x = bedrooms, y = price, color = bathrooms)) + 
    geom_point(position = "jitter") + 
    scale_y_continuous(labels = dollar) + 
    geom_smooth(method = "lm", se = FALSE),
  
  # price, grade, yr_renovated
  zillow %>% 
    filter(yr_renovated > 1990) %>% 
    mutate(grade = as.factor(grade)) %>% 
    ggplot(mapping = aes(x = yr_renovated, y = price, color = grade)) + 
    geom_point(position = "jitter") + 
    scale_y_continuous(labels = dollar) + 
    geom_smooth(method = "lm", se = FALSE)
    
)
```

**Comments**

+ Similar relationship between grade and price and bedrooms and price

    - can tell by similar slope lines
    - filtered bedrooms to focus on bedrooms < 5
    - filtered bathrooms to focus on bathrooms < 4
    
+ yr_renovated and grade don't drive price as I expected it would
    

**Questions**

+ Try faceted box plots

```{r}
grid.arrange(
  # grade vs price
  zillow %>% 
    mutate(grade = as.factor(grade)) %>% 
    ggplot(mapping = aes(x = grade, y = price)) + 
    geom_boxplot() + 
    scale_y_continuous(labels = dollar), 
  
  # condition vs price
  zillow %>% 
    mutate(condition = as.factor(condition)) %>% 
    filter(price < 4000000) %>% 
    ggplot(mapping = aes(x = condition, y = price)) + 
    geom_boxplot() + 
    scale_y_continuous(labels = dollar), 
  
  
  # bathrooms vs price
  zillow %>% 
    filter(bathrooms %in% bath) %>% 
    mutate(bathrooms = as.factor(bathrooms)) %>% 
    ggplot(mapping = aes(x = bathrooms, y = price)) + 
    geom_boxplot() + 
    scale_y_continuous(labels = dollar)
)
```

**Comments**

+ grade, condition, bathrooms vs price

    - price is most effected by grade compared to condition or bathrooms
    - price increases with increasing grade and bathrooms; makes sense
    
**Questions**

+ Why does median price drop between condition 3 and condition 4?

# Detailed EDA - questions raised in Base EDA 

## Does price differ by grade or condition?

```{r}
# Faceted scatter plot by grade
zillow %>% 
  ggplot(mapping = aes(x = grade, y = price)) + 
  geom_point(position = "jitter") + 
  facet_wrap( ~ condition) + 
  scale_y_continuous(labels = dollar)
```
 
```{r}
# Faceted scatter plot by condition
zillow %>% 
  ggplot(mapping = aes(x = condition, y = price)) + 
  geom_point(position = "jitter") + 
  facet_wrap( ~ grade) + 
  scale_y_continuous(labels = dollar)
```

 **Comments**
 
    - the pattern is consistent across conditions
    - majority of observations found in condition 3 and condition 4
    
## How does price vary by condition, grade, and bathrooms?

```{r}
# price relationship differ by condition and bathrooms
grid.arrange(
  # condition
zillow %>% 
  mutate(condition = as.factor(condition)) %>% 
  filter(price < 4000000) %>% 
  ggplot(mapping = aes(x = grade, y = price)) +
  geom_point(position = "jitter") +
  scale_y_continuous(labels = dollar) + 
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) + 
  geom_smooth(mapping = aes(color = condition), method = "lm", se = FALSE),

  # bathrooms
zillow %>% 
  filter(price < 4000000) %>% 
  filter(bathrooms %in% bath) %>% 
  mutate(bathrooms = as.factor(bathrooms)) %>% 
  ggplot(mapping = aes(x = grade, y = price)) +
  geom_point(position = "jitter") +
  scale_y_continuous(labels = dollar) + 
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) + 
  geom_smooth(mapping = aes(color = bathrooms), method = "lm", se = FALSE), 

  # sqft_living
zillow %>% 
  filter(price < 4000000) %>% 
  filter(sqft_living < 6000) %>% 
  filter(bathrooms %in% bath) %>% 
  mutate(bathrooms = as.factor(bathrooms)) %>% 
  ggplot(mapping = aes(x = sqft_living, y = price)) +
  geom_point(position = "jitter") +
  scale_y_continuous(labels = dollar) + 
  geom_smooth(method = "lm", se = FALSE)

)
```

**Comments - condition**

    - condition intersects for all conditions at grade 7
        - homes with sub grade 7 rating has odd distribution
        - likely due to counts of homes with higher condition being low for lower grade homes
        - counts of grade 3 homes with condition 5 are low; this makes sense
    - higher grade, higher price; makes sense
    - higher condition, steeper slope
    - price has positive correlation with both grade and condition

**Comments - bathrooms**

    - 4 bathroom houses are most expensive; makes sense
    - 4 bathroom homes have higher price points with higher grades
        - if sellers can improve the grade of their homes, they will be able to sell at much higher prices, even at grade 7
    - even a grade 9 homes with 1 bathroom won't sell for much. might be worth looking into renovating additional bathrooms to the house
    - grade 6 homes with 3.5 bathrooms sell lower than grade 6 homes with 3 bathrooms
        - values of 3.5 bathroom home isn't valued higher than 3 bathroom until it reaches grade 10
        - market seems to value 3 bathroom homes more than 3.5 bathrooms unless the 3.5 bathroom home has higher grade

```{r}
# price relationship differ -  bathrooms
vis_01 <- zillow %>% 
  filter(price < 4000000) %>% 
  filter(bathrooms %in% bath) %>% 
  mutate(bathrooms = as.factor(bathrooms)) %>% 
  ggplot(mapping = aes(x = grade, y = price)) +
  geom_point(position = "jitter") +
  scale_y_continuous(labels = dollar) + 
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) + 
  geom_smooth(mapping = aes(color = bathrooms), method = "lm", se = FALSE) + 
  labs(x = "Grade", 
       y = "Price", 
       title = "Home pricing trends with respect to Grades and Bathrooms")

vis_01

# save this visual for memo
ggsave(filename = "vis_01.png", plot = vis_01)
```


## What is relationship between grade and yr_renovated?

```{r}
grid.arrange(
zillow %>% 
  filter(yr_renovated > 1975) %>% 
  mutate(grade = as.factor(grade)) %>% 
  ggplot(mapping = aes(x = yr_renovated, fill = grade)) + 
  geom_bar(), 

# proportion of counts
zillow %>% 
  filter(yr_renovated > 1975) %>% 
  mutate(grade = as.factor(grade)) %>%
  ggplot(mapping = aes(x = yr_renovated, fill = grade)) + 
  geom_bar(position = "fill")
)
  
```

```{r}
# proportion of counts
vis_02 <- zillow %>% 
  filter(yr_renovated > 1975) %>% 
  mutate(grade = as.factor(grade)) %>%
  ggplot(mapping = aes(x = yr_renovated, fill = grade)) + 
  geom_bar(position = "fill") + 
  labs(x = "Year Renovated", 
       y = "Proportion", 
       title = "Proportion or grade rating with year renovated")

vis_02

# save this visual for memo
ggsave(filename = "vis_02.png", plot = vis_02)
```


**Comments**

    - Renovations seem to be made to achieve grade 7, 8 or 9

**Questions**

+ Why aren't renovations being made to achieve higher than grade 9 rating? This would increase property value
    
## Are the condition = 5 homes brand new built homes or newly renovated homes?
    
```{r}
grid.arrange(
  
zillow %>% 
  filter(yr_built > 1920) %>% 
  ggplot(mapping = aes(x = condition, y = yr_built)) + 
    geom_point(position = "jitter"), 

zillow %>% 
  filter(yr_renovated > 1920) %>% 
  ggplot(mapping = aes(x = condition, y = yr_renovated)) + 
    geom_point(position = "jitter")
  )
  
```

**Comments**

    - Majority of condition 5 homes were built that way, not renovated that way
    
**Questions**

+ How does price compare between yr_built and yr_renovated?

```{r}
grid.arrange(
  # Look at median price over time
  zillow %>% 
    filter(yr_built > 1920) %>% 
    group_by(yr_built) %>% 
    summarize(med_price = median(price)) %>% 
    filter(med_price < 900000) %>% 
    ggplot(mapping = aes(x = yr_built, y = med_price)) + 
    geom_line() + 
    scale_y_continuous(labels = dollar, limits = c(300000, 900000)),
  
  zillow %>% 
    filter(yr_renovated > 1920) %>%   # remove outliers and follow same timeline as yr_built
    group_by(yr_renovated) %>% 
    summarize(med_price = median(price)) %>% 
    ggplot(mapping = aes(x = yr_renovated, y = med_price)) + 
    geom_line() + 
    scale_y_continuous(labels = dollar, limits = c(300000, 900000))
    
)
```

**Comments**

+ In 2010, price it was more expensive to buy a newly renovated home than it was to buy a newly built home

    - In 2015, median price for a newly built home vs a newly renovated home was relatively the same
    
+ Interesting that price is similar in 2015 between both yr_built and yr_renovated. 

    - I would have assumed that newly built would always be higher price point than newly renovated
    - Houses aren't flipping for much. House flippers couldn't beat the market in 2015
    - Houses could be flipped for a profit between 1990 - ~2008
    
+ More spikes in price when looking at yr_renovated compared to yr_built

    - 90s and 00s had higher prices for yr_renovated
        - This indicates sellers markets to me - buy low, sell high
    
+ gaps in yr_renovated data in 50s, 60s and 80s

**Questions**

+ No further questions


# Summary

After extensive exploration of this dataset, the following key takeaways have been learned

1. The King County housing market is flooded with homes with grade ratings of grade 7, grade 8, and grade 9. 

2. Grade was found to be the second highest correlated variable to price, highest being sqft_living. 

3. Number of bathrooms was found to be the third highest correlated variable to price

4. Renovations have increased in popularity from 1980 onward. 

    - Depending on housing market conditions, newly renovated homes are valued higher than newly built homes 

5. Renovations appear to achieve grade 7, grade 8 or grade 9. Sellers should focus on achieving higher than grade 9 ratings to increase their property value. 

    - Home owners with more bathrooms and a higher grade home have highly valued homes

6. Potential further research: 

    - Have home renovations increased due to the pandemic?
    - Have sales at Home Depot, Lowes, and other DIY retail increased due to the pandemic?
    - Have sales of contractor services increased due to the pandemic?
    - Have sales of raw construction materials increased due to the pandemic?


